---
title: "qx2277_hw2_p8105"
author: "Qiaoen Xia"
date: "`r Sys.Date()`"
output: github_document
---

```{r}
library("tidyverse")
library(readr)
```
#Q1
```{r}
NYC_Data <- read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
head(NYC_Data)
variable.names(NYC_Data)
head(NYC_Data)
```

```{R}
NYC_clean = NYC_Data |>
  select(
    line = Line,
    station_name = `Station Name`,
    station_latitude = `Station Latitude`,
    station_longitude = `Station Longitude`,
    route1 = Route1,
    route2 = Route2,
    route3 = Route3,
    route4 = Route4,
    route5 = Route5,
    route6 = Route6,
    route7 = Route7,
    route8 = Route8,
    route9 = Route9,
    route10 = Route10,
    route11 = Route11,
    entry = Entry,
    vending = Vending,
    entrance_type = `Entrance Type`,
    ada = ADA
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

glimpse(NYC_clean)
dim(NYC_clean)
class(NYC_clean$entry)
```

```{r}
sum(duplicated(NYC_clean)) #1184

duplicate_rows = NYC_clean[
  duplicated(NYC_clean) |
    duplicated(NYC_clean, fromLast = TRUE),
]

#View(duplicate_rows)#1641



#- 1,641 rows: all records belonging to repeated groups, including the first occurrence in each group.
#- 1,184 rows: additional records after the first occurrence in each group.
#- 1,641 − 1,184 = 457: 457 distinct row combinations appeared more than once.
#- 1,868 − 1,641 = 227: 227 distinct row combinations appeared exactly once.


row_counts = count(
  NYC_clean,
  across(everything()),
  name = "n_copies",
  sort = TRUE
)


repeated_groups = filter(row_counts, n_copies > 1)

head(repeated_groups)#showing how many copies in the last column
```
The dataset contains information on NYC subway entrances and exits, including subway line, station name and coordinates, routes served, entry permission, vending availability, entrance type, and ADA compliance. I retained these variables, manually renamed the columns, and converted entry from “YES”/“NO” to a logical variable. The resulting dataset contains 1,868 rows and 19 columns. I examined repeated row combinations and found 684 distinct combinations: 227 appeared once, while 457 appeared multiple times, accounting for 1,641 records and 1,184 additional copies beyond the first occurrence. I created an n_copies variable to examine these frequencies and used distinct() to create a separate dataset containing 684 rows and 19 columns.
```{r}
NYC_unique = distinct(NYC_clean)


dim(NYC_unique)
```

```{r}
NYC_stations = distinct(
  NYC_unique,
  line,
  station_name
)

nrow(NYC_stations)
```

```{r}
table(NYC_unique$ada, useNA = "ifany")

ADA_records = filter(NYC_unique, ada == TRUE)

ADA_stations = distinct(
  ADA_records,
  line,
  station_name
)

nrow(ADA_stations)
```



#Q2
```{r}
library(readxl)

Mr_Trash_Wheel_dat <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel")
head(Mr_Trash_Wheel_dat)


Mr_Trash_Wheel_clean = Mr_Trash_Wheel_dat[, 1:14]

Mr_Trash_Wheel_clean = Mr_Trash_Wheel_clean[
  -nrow(Mr_Trash_Wheel_clean),
]


dim(Mr_Trash_Wheel_clean)
tail(Mr_Trash_Wheel_clean)

names(Mr_Trash_Wheel_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "glass_bottles",
  "plastic_bags",
  "wrappers",
  "sports_balls",
  "homes_powered"
)

names(Mr_Trash_Wheel_clean)

Mr_Trash_Wheel_clean$sports_balls = as.integer(
  round(Mr_Trash_Wheel_clean$sports_balls)
  
)

head(Mr_Trash_Wheel_clean)
```

```{r}
Professor_Trash_Wheel_dat <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel")

Professor_Trash_Wheel_clean = Professor_Trash_Wheel_dat[
  -nrow(Professor_Trash_Wheel_dat),
]

tail(Professor_Trash_Wheel_clean)

variable.names(Professor_Trash_Wheel_clean)

names(Professor_Trash_Wheel_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "glass_bottles",
  "plastic_bags",
  "wrappers",
  "homes_powered"
)

Professor_Trash_Wheel_clean$plastic_bottles = as.numeric(
  Professor_Trash_Wheel_clean$plastic_bottles
)

class(Professor_Trash_Wheel_clean$plastic_bottles)

```

```{r}
Gwynnda_dat <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda the Good Wheel of the W")
head(Gwynnda_dat)

Gwynnda_clean = filter(
  Gwynnda_dat,
  !is.na(Dumpster)
)

tail(Gwynnda_clean)
dim(Gwynnda_clean)

variable.names(Gwynnda_clean)
names(Gwynnda_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "plastic_bags",
  "wrappers",
  "homes_powered"
)

```
```{r}
sum(duplicated(Mr_Trash_Wheel_clean))

sum(duplicated(Professor_Trash_Wheel_clean))

sum(duplicated(Gwynnda_clean))
```

```{r}
Mr_Trash_Wheel_clean$trash_wheel = "Mr. Trash Wheel"
Professor_Trash_Wheel_clean$trash_wheel = "Professor Trash Wheel"
Gwynnda_clean$trash_wheel = "Gwynnda"


```

```{r}
Mr_Trash_Wheel_clean$trash_wheel = "Mr. Trash Wheel"
Professor_Trash_Wheel_clean$trash_wheel = "Professor Trash Wheel"
Gwynnda_clean$trash_wheel = "Gwynnda"

Mr_Trash_Wheel_clean$year = as.numeric(
  Mr_Trash_Wheel_clean$year
)#previously class char

class(Mr_Trash_Wheel_clean$year)
```

```{r}
trash_wheel_combined = bind_rows(
  Mr_Trash_Wheel_clean,
  Professor_Trash_Wheel_clean,
  Gwynnda_clean
)

dim(trash_wheel_combined)
```


```{r}
professor_data = filter(
  trash_wheel_combined,
  trash_wheel == "Professor Trash Wheel"
)

professor_total_weight = sum(
  professor_data$weight_tons,
  na.rm = TRUE
)

gwynnda_june_2022 = filter(
  trash_wheel_combined,
  trash_wheel == "Gwynnda",
  year == 2022,
  month == "June"
)

gwynnda_total_cigarettes = sum(
  gwynnda_june_2022$cigarette_butts,
  na.rm = TRUE
)




professor_total_weight
gwynnda_total_cigarettes
```

The combined Trash Wheel dataset contains 1188 observations and 15 variables. Each observation represents a dumpster collection record from Mr. Trash Wheel, Professor Trash Wheel, or Gwynnda. Key variables include collection date, trash weight in tons, volume in cubic yards, and counts of items such as plastic bottles, cigarette butts, and wrappers. I removed non-data entries, mostly last row, standardized variable names, converted year to a consistent numeric type, and rounded sports ball counts to integers. for Professor trash, I converted the class of the variable plastic_bottles to numeric.I added a trash wheel identifier before combining the datasets; variables unavailable for a particular trash wheel were filled with NA. Based on available data, Professor Trash Wheel collected 282.26 tons of trash, and Gwynnda collected 18120 cigarette butts in June 2022.









#Q3
```{r}
mci_amyloid <- read_csv("data/mci_amyloid.csv", skip = 1, na = c("", "NA", "Na", "na", "N/A", "."))
head(mci_amyloid)


MCI_baseline = read_csv(
  "data/MCI_baseline.csv",
  skip = 1, na = c("", "NA", "Na", "na", "N/A", ".")
)



variable.names(mci_amyloid)
variable.names(MCI_baseline)


```


```{r}
names(MCI_baseline) = c(
  "id",
  "baseline_age",
  "sex",
  "education",
  "apoe4",
  "mci_onset_age"
)



MCI_baseline$sex = factor(
  MCI_baseline$sex,
  levels = c(0, 1),
  labels = c("Female", "Male")
)

MCI_baseline$apoe4 = factor(
  MCI_baseline$apoe4,
  levels = c(0, 1),
  labels = c("Non-carrier", "Carrier")
)

```


```{r}
MCI_baseline$mci_onset_age = as.numeric(
  MCI_baseline$mci_onset_age
)

MCI_baseline_clean = filter(
  MCI_baseline,
  is.na(mci_onset_age) | mci_onset_age > baseline_age
)

nrow(MCI_baseline_clean)

```
```{r}
MCI_developed = filter(
  MCI_baseline,
  !is.na(mci_onset_age) & mci_onset_age > baseline_age
)

nrow(MCI_developed)
```
```{r}
average_baseline_age = mean(
  MCI_baseline_clean$baseline_age,
  na.rm = TRUE
)

average_baseline_age
```

```{r}
women = filter(
  MCI_baseline_clean,
  sex == "Female"
)

table(women$apoe4, useNA = "ifany")

prop_women_carriers = mean(
  women$apoe4 == "Carrier",
  na.rm = TRUE
)

prop_women_carriers
```

I imported the baseline dataset using skip = 1 to omit the initial description row and specified missing-value codes, including ".". I standardized variable names and converted sex and APOE4 carrier status to factors with descriptive labels. I excluded participants whose MCI onset age was less than or equal to their baseline age, while retaining participants with missing onset ages because they remained MCI-free during follow-up. The original dataset contained 483 participants, of whom 479 met the inclusion criteria. Each row in the cleaned dataset represents one participant, with variables for participant ID, baseline age, sex, education, APOE4 status, and MCI onset age. Among eligible participants, 79 developed MCI during follow-up. 


```{r}
names(mci_amyloid) = c(
  "id",
  "time_0",
  "time_2",
  "time_4",
  "time_6",
  "time_8"
)

names(mci_amyloid)

```

I imported the amyloid dataset by skipping the initial description row and specifying missing-value codes so that entries such as "NA" and "Na" were treated consistently. I renamed the columns to identify participant ID and measurement times. The dataset contains 487 participants, with amyloid 42/40 ratios recorded at baseline and at 2, 4, 6, and 8 years. The original structure was wide, with one row per participant and separate columns for each measurement time. I used pivot_longer() to reshape the matched data into a tidy long format, with separate variables for time and amyloid ratio, and converted time to numeric. Each row in the resulting dataset represents one participant at one scheduled measurement time. Missing biomarker values were retained.


```{r}

baseline_ids = unique(MCI_baseline_clean$id)
amyloid_ids = unique(mci_amyloid$id)


baseline_only = setdiff(baseline_ids, amyloid_ids)


amyloid_only = setdiff(amyloid_ids, baseline_ids)

baseline_only
amyloid_only


length(baseline_only)
length(amyloid_only)
```
Eight participants appear only in the cleaned baseline dataset, while 16 appear only in the amyloid dataset. Thus, the participant IDs do not completely overlap.



```{r}
mci_combined = inner_join(
  MCI_baseline_clean,
  mci_amyloid,
  by = "id"
)




mci_combined = pivot_longer(
  mci_combined,
  cols = starts_with("time_"),
  names_to = "time",
  names_prefix = "time_",
  values_to = "amyloid_ratio"
)
mci_combined$time = as.numeric(mci_combined$time)

head(mci_combined)


dim(mci_combined)

write_csv(
  mci_combined,
  "data/mci_combined.csv")
```