---
title: "qx2277_hw2_p8105"
author: "Qiaoen Xia"
date: "`r Sys.Date()`"
output: github_document
---

```{r}
library("tidyverse")
library(readr)
```
#Q1
```{r}
NYC_Data <- read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
View(NYC_Data)
variable.names(NYC_Data)
head(NYC_Data)
```

```{R}
NYC_clean = NYC_Data |>
  select(
    line = Line,
    station_name = `Station Name`,
    station_latitude = `Station Latitude`,
    station_longitude = `Station Longitude`,
    route1 = Route1,
    route2 = Route2,
    route3 = Route3,
    route4 = Route4,
    route5 = Route5,
    route6 = Route6,
    route7 = Route7,
    route8 = Route8,
    route9 = Route9,
    route10 = Route10,
    route11 = Route11,
    entry = Entry,
    vending = Vending,
    entrance_type = `Entrance Type`,
    ada = ADA
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

glimpse(NYC_clean)
dim(NYC_clean)
class(NYC_clean$entry)
```

```{r}
sum(duplicated(NYC_clean)) #1184

duplicate_rows = NYC_clean[
  duplicated(NYC_clean) |
    duplicated(NYC_clean, fromLast = TRUE),
]

View(duplicate_rows)#1641



#- 1,641 rows: all records belonging to repeated groups, including the first occurrence in each group.
#- 1,184 rows: additional records after the first occurrence in each group.
#- 1,641 − 1,184 = 457: 457 distinct row combinations appeared more than once.
#- 1,868 − 1,641 = 227: 227 distinct row combinations appeared exactly once.


row_counts = count(
  NYC_clean,
  across(everything()),
  name = "n_copies",
  sort = TRUE
)

# 只查看出现超过一次的组合
repeated_groups = filter(row_counts, n_copies > 1)

View(repeated_groups)
```
The dataset contains information on NYC subway entrances and exits, including subway line, station name and coordinates, routes served, entry permission, vending availability, entrance type, and ADA compliance. I retained these variables, manually renamed the columns, and converted entry from “YES”/“NO” to a logical variable. The resulting dataset contains 1,868 rows and 19 columns. I examined repeated row combinations and found 684 distinct combinations: 227 appeared once, while 457 appeared multiple times, accounting for 1,641 records and 1,184 additional copies beyond the first occurrence. I created an n_copies variable to examine these frequencies and used distinct() to create a separate dataset containing 684 rows and 19 columns.
```{r}
NYC_unique = distinct(NYC_clean)


dim(NYC_unique)
```

```{r}
NYC_stations = distinct(
  NYC_unique,
  line,
  station_name
)

nrow(NYC_stations)
```

```{r}
table(NYC_unique$ada, useNA = "ifany")

ADA_records = filter(NYC_unique, ada == TRUE)

ADA_stations = distinct(
  ADA_records,
  line,
  station_name
)

nrow(ADA_stations)
```



#Q2
```{r}
library(readxl)

Mr_Trash_Wheel_dat <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel")
View(Mr_Trash_Wheel_dat)


Mr_Trash_Wheel_clean = Mr_Trash_Wheel_dat[, 1:14]

Mr_Trash_Wheel_clean = Mr_Trash_Wheel_clean[
  -nrow(Mr_Trash_Wheel_clean),
]


dim(Mr_Trash_Wheel_clean)
view(Mr_Trash_Wheel_clean)
tail(Mr_Trash_Wheel_clean)

names(Mr_Trash_Wheel_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "glass_bottles",
  "plastic_bags",
  "wrappers",
  "sports_balls",
  "homes_powered"
)

names(Mr_Trash_Wheel_clean)

Mr_Trash_Wheel_clean$sports_balls = as.integer(
  round(Mr_Trash_Wheel_clean$sports_balls)
  
)

view(Mr_Trash_Wheel_clean)
```

```{r}
Professor_Trash_Wheel_dat <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel")

Professor_Trash_Wheel_clean = Professor_Trash_Wheel_dat[
  -nrow(Professor_Trash_Wheel_dat),
]

tail(Professor_Trash_Wheel_clean)

variable.names(Professor_Trash_Wheel_clean)

names(Professor_Trash_Wheel_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "glass_bottles",
  "plastic_bags",
  "wrappers",
  "homes_powered"
)


```

```{r}
Gwynnda_dat <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel")
View(Gwynnda_dat)

Gwynnda_clean = filter(
  Gwynnda_dat,
  !is.na(Dumpster)
)

tail(Gwynnda_clean)
dim(Gwynnda_clean)

variable.names(Gwynnda_clean)
names(Gwynnda_clean) = c(
  "dumpster",
  "month",
  "year",
  "date",
  "weight_tons",
  "volume_cubic_yards",
  "plastic_bottles",
  "polystyrene",
  "cigarette_butts",
  "plastic_bags",
  "wrappers",
  "homes_powered"
)

```
```{r}
sum(duplicated(Mr_Trash_Wheel_clean))

sum(duplicated(Professor_Trash_Wheel_clean))

sum(duplicated(Gwynnda_clean))
```

```{r}
Mr_Trash_Wheel_clean$trash_wheel = "Mr. Trash Wheel"
Professor_Trash_Wheel_clean$trash_wheel = "Professor Trash Wheel"
Gwynnda_clean$trash_wheel = "Gwynnda"


```

```{r}
Mr_Trash_Wheel_clean$trash_wheel = "Mr. Trash Wheel"
Professor_Trash_Wheel_clean$trash_wheel = "Professor Trash Wheel"
Gwynnda_clean$trash_wheel = "Gwynnda"

Mr_Trash_Wheel_clean$year = as.numeric(
  Mr_Trash_Wheel_clean$year
)#previously class char

class(Mr_Trash_Wheel_clean$year)
```

```{r}
trash_wheel_combined = bind_rows(
  Mr_Trash_Wheel_clean,
  Professor_Trash_Wheel_clean,
  Gwynnda_clean
)

dim(trash_wheel_combined)
```


```{r}
professor_data = filter(
  trash_wheel_combined,
  trash_wheel == "Professor Trash Wheel"
)

professor_total_weight = sum(
  professor_data$weight_tons,
  na.rm = TRUE
)

gwynnda_june_2022 = filter(
  trash_wheel_combined,
  trash_wheel == "Gwynnda",
  year == 2022,
  month == "June"
)

gwynnda_total_cigarettes = sum(
  gwynnda_june_2022$cigarette_butts,
  na.rm = TRUE
)




professor_total_weight
gwynnda_total_cigarettes
```